Implications of afforestation policies for global markets

### 2024-02-27 paper.Rmd
Update paper.Rmd

Steinbuks and Hertel [@steinbuks2012] developed the FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) as a global partial equilibrium model to analyze optimal land use within a context of increasing demand for food, bioenergy, forest products, and non-provisioning ecosystem services, while also meeting greenhouse gas targets. The model encompasses nine sectors, including agriculture, livestock, food processing, biofuels, energy, forest, timber processing, and ecosystem services. While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it primarily focuses on specific sectors of the economy, neglecting general equilibrium effects and socio-economic or household welfare impacts [@steinbuks2012] [@hertel2017]. Other efforts have used the Global Trade Analysis Project (GTAP) computable general equilibrium model to address such questions, but its standard model does not well account for land-use changes. As a result, the Land Use and Land Cover Database within the GTAP framework, which incorporates forestry remote-sensing products, was created [@baldos2012]. It was further strengthened as part of the GTAP Agro-Ecological Zone (GTAP-AEZ) model, with the use of spatially explicit global land use data and through the incorporation of intra- and inter-regional land and land-based greenhouse emissions heterogeneity [@hertel2008]. While the GTAP-AEZ model benefits from the inclusion of competition among different crops, grazing, and forest-based uses, it has limitations, primarily due to its infrequent updates. The KLUM@GTAP framework integrated the Kleines Land Use Model (KLUM) with an extended yet static version of GTAP called GTAP-EFL to evaluate the impacts of climate change on cropland allocation. GTAP-EFL separates energy factors from intermediate inputs and incorporates them into capital, while also considering CO2 emissions. KLUM, a global agricultural land-use model, connects the economy to global crop allocation to maximize producer returns under specific risk assumptions. KLUM@GTAP substitutes the land allocation mechanism within GTAP-EFL by utilizing regionally aggregated area changes in cropland determined by KLUM to update cropland shares in GTAP-EFL [@ronneberger2009].

More recent modeling attempts of land-use change have begun to move away from comparative static analysis. For example, the dynamic GTAP model, supported by the GTAP-AEZ database, considers land market effects, which were identified as significant in driving results by Stevenson et al. [@stevenson2013], and has been utilized to evaluate the impacts of crop intensification on land use. It, however, has been critiqued, as global aggregates may mask localized shifts that can have implications for ecologically significant areas [@byerlee2014]. In addition, in the standard GTAP model, it is difficult to model a sector individually. There are 68 market sectors covered in the latest version of the GTAP database. The GTAP model's general equilibrium nature means that all these markets are cleared with market-clearing prices. As the model applies nested CES production functions to all sectors and differentiates them by parameters only, to model one sector, in this case forestry, differently from the rest would require a major redesign of the model. While not impossible, this project will attempt to get around these concerns and assess the impact of various likely land-use scenarios by melding GTAP with the Global Timber Model (GTM), which is a dynamic optimization model that has been specifically designed to analyze the relationship among land rents, forested land cover, and carbon sequestration through forests and thus is better able to examine the impact of afforestation or avoided deforestation on the carbon sink [@sohngen2001]. The latest versions of both models were used. For the GTAP model, this meant using the 11b data which has 2017 as a base year. This data was then updated to the base year (2023) using data from the Centre for Prospective Studies and International Information (CEPII) to account for the projected changes in gross domestic product (GDP), population, and productivity for each country in the world.

>>>>

Toward this end, Steinbuks and Hertel [@steinbuks2012] developed their global partial equilibrium FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) to identify the optimal land use choices given the competing demands of meeting greenhouse gas targets and satisfying demand for food, bioenergy, forest products, and ecosystem services. While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it is focused on specific sectors of the economy and neglects general equilibrium and household welfare effects [@steinbuks2012] [@hertel2017]. Other efforts have used the Global Trade Analysis Project (GTAP) computable general equilibrium model to address such questions, but its standard model does not well account for land-use changes. This prompted the creation of the Land Use and Land Cover Database within the GTAP framework, which incorporates forestry remote-sensing products [@baldos2012]. The GTAP Agro-Ecological Zone (GTAP-AEZ) model further strengthened the GTAP model's ability to analyze land use changes, with the use of spatially explicit global land use data and through the incorporation of intra- and inter-regional land and land-based greenhouse emissions heterogeneity [@hertel2008]. While the GTAP-AEZ model benefits from the inclusion of competition among different crops, grazing, and forest-based uses, it has limitations, primarily due to its infrequent updates. The KLUM@GTAP framework integrated the Kleines Land Use Model (KLUM) with an extended yet static version of GTAP called GTAP-EFL to evaluate the impact of climate change on cropland allocation. GTAP-EFL separates energy factors from intermediate inputs and incorporates them into capital, while also considering CO2 emissions. KLUM, a global agricultural land-use model, connects the economy to global crop allocation to maximize producer returns under specific risk assumptions. KLUM@GTAP substitutes the land allocation mechanism within GTAP-EFL by utilizing regionally aggregated area changes in cropland determined by KLUM to update cropland shares in GTAP-EFL [@ronneberger2009].

More recent modeling attempts of land-use change have begun to move away from comparative static analysis. For example, the dynamic GTAP model, supported by the GTAP-AEZ database, considers land market effects, which were identified as significant in driving results by Stevenson et al. [@stevenson2013]. This model has been utilized to evaluate the impact of crop intensification on land use. It, however, has been critiqued, as global aggregates may mask localized shifts that can have implications for ecologically significant areas [@byerlee2014]. In addition, in the standard GTAP model, it is difficult to model a sector individually. There are 68 market sectors covered in the latest version of the GTAP database. The GTAP model's general equilibrium nature means that all these markets are cleared with market-clearing prices. As the model applies nested CES production functions to all sectors and differentiates them by parameters only, to model one sector, in this case forestry, differently from the rest would require a major redesign of the model. While not impossible, this project will attempt to get around these concerns and assess the impact of various likely land-use scenarios by melding the dynamic GTAP model with the Global Timber Model (GTM), which is a dynamic optimization model that has been specifically designed to analyze the relationship among land rents, forested land cover, and carbon sequestration through forests and thus is better able to examine the impact of afforestation or avoided deforestation on the carbon sink [@sohngen2001]. The latest versions of both models were used. For the GTAP model, this meant using data which has 2017 as its base year. This data was then updated to the base year (2023) using data from the Centre for Prospective Studies and International Information (CEPII) to account for the projected changes in gross domestic product (GDP), population, and productivity for each country in the world. Similar steps were taken to update and recalibrate the GTM model.
